\input{preamble}
\input{format}
\input{commands}

\begin{document}

\textsf{\LARGE{\textbf{Take-home exam 2}}}

\normalsize{\textit{Dynamical Systems}}

\vspace{1ex}

\textsf{\textbf{Student:}} \text{Males-Araujo Yorlan}, 
\href{mailto:yorlan.males@yachaytech.edu.ec}{\texttt{yorlan.males@yachaytech.edu.ec}}\\
\textsf{\textbf{Lecturer:}} \text{Mario Cosenza}, 
\href{mcosenza@yachaytech.edu.ec}{\texttt{mcosenza@yachaytech.edu.ec}}

Yachay Tech, \today.

\vspace{2ex}



\begin{problem}{Superestable orbit}{problem-label-2}

Let $r_n^*$ be the parameter value of a quadratic map $f_r(x)$ for which there
exists a superstable orbit of period $2^n$.

\begin{enumerate}[(a)]
    \item Consider an arbitrary function $h(r_n)$ evaluated at $r_n^*$. Show that, 
    for $n \gg 1$, $h(r_n)$ satisfies the relation [1]:
    \[
        \frac{[h(r_n^*)-h(r_{\infty})]\delta^n}{h'(r_{\infty})} = \text{cte.}
    \]
    \item Show that for $ n \gg 1$, [1]
    \[
        (r_n^*-r_{\infty})\delta^n \propto \frac{\delta^2}{\delta -1}.
    \]
\end{enumerate}
\end{problem}

Solution here.

\begin{problem}{Schwarz derivative}{problem-label-3}

Consider the logarithmic map $x_{n+1}=f(x_n)=b+\log|x_n|$.

\begin{enumerate}[(a)]
    \item Calculate the Schwarz derivative of this map. [1]
    \item Show that the Lyapunov exponent for this map can be expressed as [1]
    \[
        \lambda = b - \frac{1}{n}\sum_{j=1}^{n}x_j.
    \] 
\end{enumerate}
\end{problem}

Solution here.

\begin{problem}{Bifurcation diagram}{problem-label-4}

Consider the following map for $x_n \in [0,1]$, depending on two real parameters $s$ and $c$,
\[
    x_{n+1} = f(x_n) = |\tanh[s(x_n-c)]|.
\]
Obtain the bifurcation diagram of $x_n$ as a function of $c \in [0,1]$ with fixed value $s=1.3$. [1]

\end{problem}

Solution here.

\begin{problem}{Lyapunov exponent}{problem-label-5}

Consider the following map for $x_n \in [0,1]$:
\[
    x_{n+1} = f(x_n) = \begin{cases}
        (1+\epsilon)x_n + x_n^2, & \text{if } x_n < x^*, \\
        1-x_n, & \text{if } x_n \geq x^*,
    \end{cases}
\]
where $x^*$ is the solution of $(1+\epsilon)x_n + x_n^2 = 1$.

\begin{enumerate}[(a)]
    \item Calculate the Lyapunov exponent versus $\epsilon \in [-0.4, 0.4]$. [1]
    \item Describe the route to chaos exhibited by this map on such interval. [1]

\end{enumerate}
\end{problem}

Solution here.

\begin{problem}{Not period-doubling}{problem-label-6}
Consider the following map for $x_n \in [0,1],$
\[
    x_{n+1} = f(x_n) = \frac{1-b^{x_n(1-x_n)}}{1-b^{1/4}}
\]
\begin{enumerate}[(a)]
    \item Is this map unimodal for $b \in (0,1)$? [1]
    \item Plot the bifurcation diagram of $x_n$ on the interval $b \in (0,1)$ and $x_n \in [0,1]$.
    Why does this map not show period doubling? [1]
    \item Find the Lyapunov exponent for $b \in (0,1)$. [1]
\end{enumerate}

Solution here.

\end{problem}

\newpage
\vspace{0.1ex}

\bibliographystyle{apalike}
\bibliography{references}

\end{document}